% This dissertation is the product of 
%     kindness and generosity invested in me by  
%
%%%%%%%%%%%%%%%%% ADVISOR %%%%%%%%%%%%%%%%
%
None of this work presented here would have been possible without the unique perspective and unwavering support of my advisor, Joe Halpern. 
Joe has reliably given me a rare combination of resources:
% (1) 
complete freedom to formulate my own research questions,
% (2) 
inhumanly fast and thorough responses to drafts and email---even the most monstrous ones (of which there were many),
%  consisting of dozens of heavy technical arguments and sharply worded disagreement, 
% (3) 
funding, and a solid assurance that support is not contingent on research output or personal disagreements (of which there were also many).
%
% Advising me cannot have been easy.
% Joe has taught me a lot about writing and editing. 
%
% I have never met someone with whom I am so entirely out of sync.
% Never before have I met someone with respect to whom I am so precisely the same wavelength yet completely out of phase. 
% It is no secret that Joe and I have had significant disagreements. 
% I may never again be so in sync with another person, yet be so completely out of phase, and struggle so profoundly to communicate. 
I have never before encountered someone with whom I am so in sync but out of phase, 
    nor experienced such profound difficulty in communicating ideas to such a smart person. 
%
I can only imagine how difficult it must have been
    to advise a PhD student
    who is as prepared as I am to shoot themselves in the foot
        in service of the slightest aesthetic or epistemic truth---%
    especially for someone cut from the opposite end of the same cloth.
%
% I know for certain that the experience was frustrating for both of us
Yet despite the frustration and conflict (or perhaps because of it), I learned a great deal from Joe. 
    From the value of killing good ideas to let great ones thrive, 
        to the compact readability English-based inline math, 
        and the power of tiny words,
    Joe taught me how to deeply develop an idea and package it for others.
    % importance of killing ideas, the value of inline math, the power of tiny words
I cannot thank him enough for his tenacious engagement, unconditional personal support, and the insane number of cycles he has invested in me. 
    % am fully prepared to shoot myself in the foot in service of Truth or Beauty. 
    % am fully prepared to shoot myself in the foot in service of the smallest truth or aesthetic choice. 
% No doubt 

%
% Every good story needs a good villain; for me, that villain has been Joe. 
% Yet far from a two-dimensional cartoon villain, Joe has 
% Over the years, Joe and I have disagreed about everything. 
%
% Joe and I care about the same things,
%     but look at the world in totally different ways. 


% Joe has invested more time in me .
% well my peers often complained that they had to wait weeks or sometimes even years to receive feedback from the advisors, much to my concernment Joe would respond in a matter of hours.


%%%%%%%%%%% COMMITTEE %%%%%%%%%%%
I owe a great deal to the wonderful faculty members on my committee, 
    all of whom have provided incredibly useful feedback on this work over the years. 
Dexter Kozen has been especially kind, supportive, and generous;  
    I really enjoyed working on bialgebra, 
    and wish I had been able to more consistently create time for the projects he gave me.
Christopher De Sa 
    % joined my committee after my A-exam, 
    and I had several very interesting research discussions,
    and it was ultimately his insight that became the basis of the inference procedure in \cref{chap:infer}. 
Ziv Goldfeld has always enthusiastically supported my work, 
        which has at times kept me from losing confidence in myself. 
    When Joe was on sabbatical, Ziv welcomed me into his lab. 
% I am grateful that these genuinely impressive people have happily written letters of recommendation on my behalf. 


%%%%%%%%%%%%%%%%% COLLABORATORS %%%%%%%%%%%%%%%
% Next, I want to acknowledge my collaborators. 
% I really appreciate my collaborators 
Next, I want to acknowledge my other research collaborators.
Jialu Bao
    % for taking a huge gamble on me and spending many confused meetings.
    took a huge gamble on me, and was instrumental in developing an important theory that lurks behind some of the material in this dissertation.
    Although we didn't complete the full Mixture Languages paper this cycle,
        I am proud of our work and am excited to finish it soon!
    % I fully expect our work on Mixture Languages to 
Spencer Peters
    % has been an effective sounding board throughout, and
    generously jumped back into the world of causality to 
    % help push our paper on qualitative modeling with hypergraphs (the material in \cref{chap:QIM})) over the finish line. 
    mediate a disagreement between myself and Joe,
    significantly improving 
        the material in \cref{chap:QIM}.
% Spencer's
His endless positivity, encouragement, and drive to engage has been a highlight and inspiration for me.  


%%%%%%%%%%%%%%%%% VARSHA %%%%%%%%%%%%%%%

Another key element of this ordeal has been my good friend (and fianc\'ee) Varsha Kishore,
    who, despite her extreme and undeniable quirks,
    has successfully convinced me that she is the epitome of grounded common sense. 
Varsha: thank you 
    for being such a reliable source of pragmatic advice,
    for offering feedback and helping revise my work, 
    for bailing me out when I get overwhelmed with logistics, 
    and for making all of it an adventure.
    % and for imbuing it all with your fiery sense of joy.
%
%%%%%%%%%%%%% Critical Friends %%%%%%%%%%%%%%%%%
Certain other peers have also had an outsized role in helping me pull this together. 
%%% DIETRICH %%%
I would like to thank
Dietrich Geisler, for making sure I applied to Cornell in the first place, and for invaluable suggestions lead to a writing breakthrough in my first paper in grad school;
%%% Greg, Kate, Raunak,
Gregory Yauney, 
    % deep friendship, 
    for editing help, strong and consistent encouragement, and the impetus to acquire a bike;
    %
Sloan Nietert, for lab comradery and interesting geometry problems;
Kate Donahue, for being a seriously good friend and always reaching out;
%%% Anshuman
and Anshuman Mohan, whose unbounded generosity fed and housed me for weeks after my lease expired and I had yet to cross the finish line.  

%%%%%%%%%%%%%%%%% CORNELL CS %%%%%%%%%%%%%%%%%
% Andrew Hirsh, Mae, Michael, 
Thank you also to the broader community in Cornell Computer Science.
In my first year, I was impressed by the department culture and civic mindedness of the older students. I would especially like to thank Mae Milano, Michael Roberts, Rediet Abebe, and Andrew Hirsh for the warm welcome. 
% Tegan, Jason,  Princewell, Emily, Haobin, 
To the great friends I made in Theory Lab, especially 
    Tegan Wilson,
    Jason Gaitonde,
    Meir Fridenberg,
    Princewill Okoroafor,    
    and Emily Riu---%
        conversing with each of you (both about research and beyond) 
        has made Gates 324 a wonderful place to be. 
% Raunak Kumar, for technical engagement and emotional support;
% Alexa Vanhattum, ;
% I feel the same about all of my friends in programming languages:
I also want to thank my friends in the programming languages community, 
    especially
    Alexa Vanhattum,
    Pedro Henrique Avezedo de Amorim,
    Rachit Nigam, 
    Haobin Ni,
    and
    Eric Campbell
    for hundreds of stimulating conversations and writing advice. 
Thanks also to Katherine van Koevering, Gloire Rubambiza, Kevin Negy, and Soham Sankaran.
% theory, programming languages, and theory 
%%%%%% Some odds and ends: %%%%
    % Soham Sankaran,
    % Shreyas Malpathak,
    % Kara Fikrig,
Finally, I want to thank Becky Stewart, for 
    % the most difficult (administrative) problems in computer science. 
    % helping all of us handle all of  problems quickly and skillfully.
    skillfully helping students navigate administrative issues,
    and otherwise making sure the department runs so smoothly.
    

This work has been supported by 
% MURI grant W911NF-19-1-0217
MURI grant W911NF-19-1-0217,
ARO grant W911NF-17-1-0592,
ARO grant W911NF-22-1-0061,
and
a grant for the Center for Human-Compatible AI from Open Philanthropy.